Leaf traits and size drive differences in above-ground biomass distribution among woody plant functional types

### 2016-01-28 manuscript.Rnw
Introduction and results written as readable draft

\begin{itemize}
\item Biomass allocation is important; need to improve GVMs and make them more consistently benchmarked against global observations. 
\item Plant functional types; can expect differences in biomass partitioning and leaf area relationships. Useful from viewpoint of ecology (functional consequences of angiosperm vs. gymnosperm divide; evergreen vs. deciduous). Also useful for GVMs which use PFT for parameterization.
\item Enquist and Niklas showed differences in leaf mass at given stem mass between gymnosperms and angiosperms.Poorter2015 showed X. Reich2014 showed Y. Open questions are 1,2,3. 
\item One major gap is that these previous studies did not include estimates of whole-plant leaf area. Leaf area is relevant because of light interception. It may follow very different patterns from leaf mass. Poorter2015 speculated that differences in leaf mass fraction are related to PFT-differences in specific leaf area, but did not have the data to directly test this prediction.
\item What about climate? We know from small-scale experiments and observational studies that biomass distribution varies with climate drivers. But unclear if global patterns exist. Reich2014 showed X, consistent with glasshouse stuff from Poorter2015/2012. 
\item We have the following objectives:
\begin{enumerate}
\item Test whether MF/MS is proportional to leaf mass per area across plant functional types (as suggested by Poorter), once corrected for the expected strong effects of plant size.
\item Partition the variance of leaf/stem biomass distribution into component variables with the objective to determine which aspects of plant allometry drive differences in biomass distributions between PFTs and with plant size. 
\item Partition variance of component variables into plant height, plant functional type and broad climate variables. 
\end{enumerate}
\end{itemize}

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.

% % less focus on GVMs; focus on ecological/biological reasons to test these things
% % ... but make sense to look at PFTs though, and this is the actual reason for doing that?
% Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010,wullschleger2014}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Notes S1), these differences are poorly constrained, due to limited available data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}.
% %explain why stem/leaf biomass distribution is important here.
% In this paper we are primarily interested in the distribution of biomass between leaves and woody stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on either whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and or a mix on stand-- and individual-based measurements \citep{poorter2012, poorter2015}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Some oft-cited studies have also assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.
% Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responsiveness of biomass partitioning to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. Moreover, a recent study by \citet{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to. 
% % integrate poorter2015 and reich2014 better
% Despite considerable advances in theory underlying allometric scaling in plants \citep{enquist2003, west1999, savage2010} we do not yet have a clear understanding of potential differences between PFTs in terms of biomass partitioning. Instead, most previous work has focussed on understanding size-related shifts in biomass partitioning, as governed by constraints including hydraulic supply and mechanical stability \citep{savage2010}. Rather than to advance a specific model of biomass partitioning or allocation, our view is that a broader evidence base is needed first to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations. Here, we use a unique, new database \citep{falster2015} (Fig.~\ref{fig:figoverview}) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. A recent compilation of plant biomass data \citep{poorter2015} speculated that differences between gymnosperms and angiosperms in distribution of biomass between leaves and stems is related to differences in leaf mass per area. Here we can directly test this hypothesis as our database, unlike those of \citet{enquistniklas2002}, \citet{reich2014}, and \cite{poorter2015}, includes many observations of leaf area as well as leaf mass measured on the same plants.
% A second objective was to study relationships between biomass partitioning and climate variables or biome at a global scale. The dataset includes observations of biomass and size metrics for individual plants, compiled from 175 studies across nine vegetation types (Fig. S1), across the three major biomes (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a semi-parametric data-driven statistical model, i.e. one that does not assume a particular functional form, thereby allowing us to study PFT patterns at a common plant height. 

@@

$$D_S = D_{Sbh} H^a / (H - H_{bh})^a$$

@@

$$a = c_0H^{c_1}.$$

@@

We used generalised additive models (GAM) to capture  the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as LMF at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalised regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. All variables (except MAP and MAT) were log-transformed prior to analysis.

Variance explained by quantitative climate variables (MAP and MAT) were tested with GAMs where variables were sequentially added to the model, and the explained variance ($R^2$) calculated. For the test of biome effects on biomass partitioning, we used a linear mixed-effects model (because two factors and their interactions were tested), again with species-dataset as the random effect. We calculated the $R^2$ for linear mixed-effects models for the fixed effects only \citep{nakagawa2013}.

All analyses were conducted in R 3.2.2 \citep{R-2015}. GAMs were fitted using the \emph{mgcv} package \citep{wood2006}. The code to replicate this analysis are available at http://github.com/RemkoDuursma/baadanalysis.

@@

\begin{itemize}
\item Dataset covers climate space pretty well. Evergreen angiosperms occurred at higher MAT than deciduous angiosperms and evergreen gymnosperms but there is much overlap between PFTs. 
\item MF and MS are non-linearly related to plant height (Fig 2), as expected. Also note stem mass increases with height more quickly than leaf mass.
\item Leaf/stem biomass distribution was very different between PFTs (Fig 3a), and proportional to LMA when estimated at a common height, so that leaf area ratio did not differ between PFTs. Consistent with variance partitioning: MF/MS mostly explained by H and PFT, but LAR not much explained by PFT.
\item Partitioning MF/MS into three component variables, and partition each of the four into H, PFT and climate. Climate effects, when analyzed in addition to H and PFT, were very weak, with only LMA and AF/AS somewhat more dependent on climate, but not consistently (Fig S4).
\item We show that AF was proportional to AS (with the exception of very large trees). AF/AS was nearly unrelated to PFT or plant height.
\item Stem mass should be proportional to basal area, height, and the product of wood density and stem shape. Across all data, the relationship between MS and AS was not related to PFT, but nearly proportional to height (as expected). For larger trees, MS at a given AS*H was clearly lower for evergreen gymnosperms, reflecting differences in wood density. Again no climate effects.
\item Histograms summarize the findings, for three variables that do not, or weakly, depend on height. Large PFT effects on LMA, small ones on stem mass index, none on AF/AS. 
\item We repeat an analysis similar to Reich et al 2014 to test for climate effects on biomass partitioning when only considering the angiosperm/gymnosperm divide, and not first correcting for H as we did previously (but instead binning by H). We find that for angiosperms, MF/MS increased with MAT for all but the smallest plants. But not for gymnosperms (Fig. 7). 
\end{itemize}

@@

% Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs (notwithstanding effects of plant height on this variable), and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Notes S1), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz2012, dekauwe2014} of GVMs against empirical data.

@@

\begin{figure}[h!]
    \centering
    \ifincludefigs 
    \includegraphics[width=0.99\textwidth]{Figure7.pdf}
    \fi
    \caption{Relationships between leaf - stem biomass distribution ($M_F/M_S$) and mean annual temperature (MAT) (degree*C). Similar to Reich et al. (2014), we divided the data into five bins of plant size (bins were chosen so that each bin had equal sample size), and used only evergreen vs. deciduous plant functional types. Lines are fits from mixed-effects models (with species within study as random effect, and only MAT and evergreen/deciduous classification as fixed effects). Solid lines are significant (p < 0.05), dashed lines are not (p > 0.05). Symbols are partial residuals as calculated with the visreg package in R.}
    \label{fig:figreichclimate}
\end{figure}

>>>>

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times is a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.

One puzzling aspect of the large inter-specific variation in biomass distribution between leaves and stems is the much larger leaf mass in gymnosperms compared to angiosperms (at a given stem mass) \citep{oneilldeangelis1981,enquistniklas2002,reich2014,poorter2012, poorter2015}. \citet{poorter2012} and \citet{enquistniklas2002} attributed higher leaf-stem biomass ratio ($M_F/M_S$) in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms. This is consistent with \citet{poorter2015} who reported lower $M_F/M_S$ in deciduous compared to evergreen species.

It is yet unknown whether plant functional types (PFTs) (angiosperm vs. gymnosperm, deciduous vs. evergreen) with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Because longer LL is generally correlated with the leaf mass per area (LMA), PFTs with higher MF/MS (and thus higher LMA) do not necessarily not have larger total leaf area at a given stem mass. Some oft-cited studies have assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}, but recent large compilations on biomass distribution did not include measurements of leaf area in addition to mass to test this hypothesis \citep{poorter2012,poorter2015,reich2014}. \citet{poorter2015} advanced the hypothesis that $MF/MS$ may be nearly proportional to LMA, arguing that total leaf area is likely less variable between PFTs. Here we set out to directly test this hypothesis with a database that includes observations of leaf area as well as leaf mass measured on the same plants.

Little is known about global-scale patterns in biomass distribution in relation to climate. It can be expected that biomass distribution between leaves and stems is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responses to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. Recently, \citet{reich2014} demonstrated that $M_F/M_S$ (for whole stands, either including or excluding below-ground components) increased with mean annual temperature, after separating the data into gymnosperm and angiosperm stands, and accounting for the large size-related variation in biomass distribution. Our second objective was thus to test for climate-related variation in above-ground biomass distribution, and specifically to test whether $M_F/M_S$ increases with mean annual temperature.

Something about plant size; we need to account for it and it is always troublesome. Also it may be confounded with climate! Thus if we want to know direct climate effects it would make sense to study it in addition to plant size, pft. Also remind us that MF/MS will decrease quickly with plant size \citep{poorter2015,reich2014,enquist2002}, because of turnover differences.

We use a recently compiled database \citet{falster2015} which includes observations of biomass and size metrics for individual plants (Fig. 1), compiled from 175 studies across nine vegetation types (Fig. S1), across the three major climate zones (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height).

@@

\begin{equation}
D_S = D_{Sbh} H^a / (H - H_{bh})^a
\end{equation}

@@

\begin{equation}
a = c_0H^{c_1}
\end{equation}

@@

To better understand the PFT and climate-related variation in leaf-stem biomass distribution, it is informative to decompose it into three component variables (Eq. 1),

\begin{equation}
\frac{M_F}{M_S} = \frac{M_F}{A_F} \cdot \frac{A_F}{A_S} \cdot \frac{A_S}{M_S}
\end{equation}

where $M_F/A_F$ is the leaf mass per area (LMA), $A_F/A_S$ the ratio of plant leaf area to basal stem area, and $A_S/M_S$ the ratio of stem basal area to total stem mass (including branches). Across PFTs, $M_F/A_F$ will thus be proportional to LMA when $A_F/A_S$ and $A_S/M_S$ do not vary between PFTs. We also expect that $M_F/M_S$ will show tremendous dependence on plant size \citep{poorter2015,reich2014}. We expect that most of this size-related variation is due to variation in $A_S/M_S$, since the total stem mass should be proportional to the plant height, stem cross-sectional area, and a shape parameter \citet{chave2005}. Using variance partitioning, we quantify the contribution of plant height, plant functional type, and climate to the overall variation in $M_F/M_S$ and for all three component variables.

Because, similar to \citep{poorter2015}, none of the studied relationships were linear on a log-log scale, we did not use standard allometric equations to study relationships between size and biomass variables. Instead, we used generalised additive models (GAM), a semi-parametric modelling approach that makes no assumption about the shape of the relationships. GAMs were used to visualize the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as $M_F/M_S$ at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalised regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. All variables (except MAP and MAT) were log-transformed prior to analysis.

Variance explained by quantitative climate variables (MAP and MAT) were tested with GAMs where variables were sequentially added to the model, and the explained variance ($R^2$) calculated. We compared this approach when using linear mixed-effects models instead of GAMs (calculating the the $R^2$ for linear mixed-effects models for the fixed effects only \citep{nakagawa2013}), as well as an independent method, independent effects analysis. The three methods generally agreed on the ranking of variable important (SI Methods).

All analyses were conducted in R 3.2.3 \citep{R-2015}. GAMs were fitted using the \emph{mgcv} package \citep{wood2006}. The code to replicate this analysis (and all figures) is available at http://github.com/RemkoDuursma/baadanalysis (doi: \emph{to be added}).

@@

%with(baad_mapmat, tapply(MAT, pft, quantile, probs=c(0.25,0.75),na.rm=T))
The compiled dataset covered the global climate space well (defined by mean annual temperature, MAT; and mean annual precipitation, MAP). There was some noticeable separation between the three major PFTs in terms of climate space occupancy. Evergreen angiosperms occurred on average at a higher MAT (50\% of studies between 15.5 and 26.0 \textdegree C MAT) than deciduous angiosperms (6.0 - 12.6 \textdegree C) and evergreen gymnosperms (3.4 - 12.4 \textdegree C). Plant functional type and climate are thus at least partly confounded.

As expected, both whole-plant leaf mass ($M_F$) and above-ground woody biomass ($M_S$) showed a strong depence on plant height (Fig. 2). For both variables the relationship was highly non-linear on a log-log scale, demonstrating departure from simple allometric relationships. It is also evident that $M_S$ increases much quicker with plant height than $M_F$. As a result, the ratio $M_F/M_S$ quickly decreased with plant height (Fig. 3a), from ca. 1-2 for small saplings (< 0.1m) to less than 0.01 for large trees (>30m).

We found large differences between the three major PFTs in terms of above-ground biomass distribution (Fig. 3a). These differences were fairly consistent along the range in plant height, with a ranking of evergreen gymnosperm > evergreen angiosperm > deciduous angiosperm. Next we calculated $M_F/M_S$ at a common plant height for each of the PFTs. These three size-corrected $M_F/M_S$ values were highly significantly different between PFTs (p < 0.001), and were approximately proportional to the average LMA ($M_F/A_F$) across the PFTs (Fig. 3b). As a result, the amount of leaf area per unit woody above-ground biomass ($A_F/M_S$) did not vary between PFTs, because the differences in $M_F/A_S$ were exactly compensated by differences in LMA (Fig. 3c).

Next, we decomposed $M_F/M_S$ into three component variables with distinct biological interpretation (Eq. 3). For each of the component variables, ad well as $M_F/M_S$ and $A_F/M_S$, we estimated the variance explained by plant height, PFT and climate variables (MAT, MAP and its interaction) (Table 1). We did this by sequentially adding terms to a GAM, and calculating the $R^2$ for each fitted model. The first model included only plant height, and already explained much variation (60 - 77\%) in $M_F/M_S$, $A_F/M_S$ and $M_S/A_S$ but not LMA and $A_F/A_S$ (0 - 7 \%). When PFT was added to the model (as an intercept term as well as interaction with the smooth H term), the $R^2$ increased substantially for $M_F/M_S$ and $A_F/A_S$ but not $A_F/M_S$ or $M_S/A_S$ (showing PFT invariance for these two variables). Additional variance explained was especially large for LMA, which is consistent with large differences shown in Fig. 3b,c.

% afas
From the variance partitioning (Table 1), it is evident that $A_F/A_S$ did not vary appreciably between PFTs. Indeed, when all PFTs were plotted together, similar $A_F$ was attained at a given $A_S$ across the entire plant size range (with the exception of extremely large trees with > 1 $m^2$ stem basal area, for which little data was available) (Fig. 4). We fit a GAM of $A_F$ as a function of $A_S$ with or without PFT as another covariate (both as intercept and slope interaction), which showed PFT significantly improved model fit (likelihood ratio test, p < 0.001), however the $R^2$ increased only from 83.8 to 85.3 \%, and thus had little if any practical significance.

Likewise, the ratio of stem basal area to stem mass ($A_S/M_S$) apparently varied little between PFTs (Table 1). To illustrate this further, when we plotted $M_S$ as a function of $A_S$ it revealed a remarkably tight relationship, with little apparent difference between PFTs (Fig 5a). The only exception was for large gymnosperm trees, which consistently showed a lower $M_S$ at a given $A_S$ (Fig. 5b). We fitted a GAM of $M_S$ as a function of $A_S$ (added as a smooth term) with or without PFT as a covariate. Again PFT was formally significant (likelihood ratio test, p < 0.001), but the $R^2$ increased only marginally from 80.8 to 82.5\%, again showing little practical significance of PFT in explaining this variable.

Taken together, these results demonstrate that 1) the majority of PFT-related variation in $M_F/M_S$ results from variation in LMA (not $A_F/A_S$ or $A_S/M_S$), and 2) the majority of height-related variation in $M_F/M_S$ results from variation in $A_S/M_S$ (not LMA or $A_F/A_S$). It is however important to note that there is very large variation between individuals even after accounting for PFT (Fig. 6). There we expressed the stem mass ratio as $M_S/(H \cdot A_S)$ because we expect from basic geometry that M_S is proportional to H and A_S ($M_S = \phi \rho A_S H$, where $\phi$ a stem shape parameter, $\rho$ wood density). The histograms summarize that PFT effects are very evident in LMA (Fig. 6a), not in $A_F/A_S$, and only a bit for $M_S/A_S$, which is entirely driven by large gymnosperm trees.

Climate variables consistently added little explanatory power to the models when added to a model that included PFT and height (1 - 2.8 \%) (Table 1). Since this result is in contrast with recent finding of \citep{reich2014}, we also analyzed climate effects with methods similar to that study. When we tested for MAT effects on $M_F/M_S$ when only considering gymnosperms and angiosperms (thus grouping decidous and evergreen species), we found a significant and consistent positive relationship with MAT (Fig. S6) for angiosperms for most size classes (but never for gymnosperms). We found, however, that this was largely due to the fact that at higher MAT, an increasing proportion of angiosperms are evergreen (Fig. 1), which have higher $M_F/M_S$. When analyzed separately for deciduous and evergreen angiosperms, few and weaker significant relationships were found (Fig. S7).

@@

A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs (notwithstanding effects of plant height on this variable), and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Notes S1), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz2012, dekauwe2014} of GVMs against empirical data.

Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010,wullschleger2014}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Notes S1), these differences are poorly constrained, due to limited available data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}. Rather than to advance a specific model of biomass partitioning or allocation, our view is that a broader evidence base is needed first to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations.

@@
